### 2015-09-21 rscripts/location_clusters.Rmd
assorted updates

lls <- lapply(8:11, function(k) movMF(pos, k, list(maxiter=10000, nruns=20)))
bics <- lapply(lls, stats::BIC)

temporal_cluster <- data.table(location_id = peaks[measure == "usage"]$location_id, vMFcluster = predict(movMF(pos, 3, list(maxiter=10000, nruns=20)), pos), key="location_id")

alt_temporal_cluster <- data.table(location_id = peaks[measure == "usage"]$location_id, vMFcluster = predict(lls[[1]], pos), key="location_id")

@@

```{r, echo=F, fig.cap="Login, Logout, Usage Peak Clusters"}
ggplot(peaks[temporal_cluster]) + theme_bw() + aes(x = peak, fill=factor(vMFcluster)) + facet_grid(measure ~ .) + geom_bar(binwidth=1) + labs(fill="von Mises-Fisher\n3-Sphere\nCluster")
```

@@

Showing the power cluster breakdown
```{r}
pre_pwr_clust <- data.table(dcast(pwr_cluster, location_id ~ measure, value.var = "pwr_cat"))
pwr_multi_clusters <- setorder(pre_pwr_clust[,.N,by=list(login, logout, usage)], -N)
pwr_multi_clusters
reduced_pwr_clust <- pre_pwr_clust[, list(pwr_clust = ifelse((login == logout) || (login == usage), login, logout)), keyby=location_id]
```
Probably worth having all-(low|med|high) clusters; the other cases all (with one exception) are two the same, the third off by one.

@@

Finally, we can look at the relative usage per user:
\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="Locations clustered excess usage per user."}
agg.dt <- initial_digest$raw[,list(log10_total_duration,log10_unique_users,log10_lifetime), keyby=location_id]
lg10lm <- lm(log10_total_duration-log10_unique_users ~ log10_lifetime, agg.dt)
agg.dt <- cbind(agg.dt, predict(lg10lm, agg.dt, interval="prediction", level=1/3))[,
  lifetime_cat := ifelse(log10_total_duration-log10_unique_users < lwr, "low",
                  ifelse(log10_total_duration-log10_unique_users < upr, "mid", "high"))
][reduced_pwr_clust][temporal_cluster]
p <- ggplot(agg.dt) + theme_bw() +
  aes(y = log10_total_duration-log10_unique_users, x = log10_lifetime, color = factor(vMFcluster), size = pwr_clust, alpha = pwr_clust) +
  with(agg.dt, annotate("ribbon", x=log10_lifetime, ymax=upr, ymin=lwr, fill="lightgrey")) +
  geom_vline(xintercept=log10(365*24*60*60), color="grey") +
  annotate("text", x=log10(365*24*60*60)+c(-0.02,0.02), y=5.75, label=c("< 1 year","> 1 year"), angle=90, color="grey") +
  geom_point() +
#  geom_ribbon(aes(ymax=upr, ymin=lwr, y=NULL, color=NULL, size=NULL), alpha=0.2) +
  labs(
    x = expression(log[10]("lifetime")),
    y = expression(log[10]*bgroup("(",frac("total usage","# unique users"),")")),
    color = "von Mises-Fisher\nPeak Time\nCluster"
  ) + scale_size_manual("power\ncluster", limits=c("low","med","hi"), values=c(2,5,8)) +
  scale_alpha_manual("power\ncluster", limits=c("low","med","hi"), values=(2^0.5)/c(2,5,8)^0.5)
p
```
\elandscape\newpage

@@

categorizer_help <- function(opts) function(SD) SD[,
  c(list(tot_visits = sum(lengths), mn_reps=mean(lengths), switches=.N),
    lapply(opts, function(v) SD[values == v, sum(lengths)])
  )
]

categorizer <- function(src.dt, cat, ref) {
  rle_expr <- paste0("rle(",cat,")")
  key_expr <- c("user_id",cat)
  mx_expr <- parse(text=paste0("list(main=",cat,"[which.max(prop)],share=max(prop))"))
  rle_dt <- src.dt[, rle(eval(parse(text=cat))), keyby=user_id]
  prop_ref <- src.dt[,list(prop=sum(logout-login)), keyby=key_expr][ref][,list(prop=prop/usage), keyby=key_expr]
  prop_lab <- prop_ref[, eval(mx_expr), by=user_id]
  chunks <- data.table(dcast(rle_dt, user_id ~ values, value.var="lengths", fun.aggregate = sum), key="user_id")
  ## cast data, maybe?
  digest <- rle_dt[, list(tot_visits = sum(lengths), mn_reps=mean(lengths), switches=.N), keyby=user_id][ref][prop_lab][chunks]
  list(rle_dt=rle_dt, prop_ref=prop_ref, prop_lab=prop_lab, digest=digest)
}

@@

\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="distribution of main location type usage"}
digests <- rbind(
  lifetime_dts$digest[,list(switch_rate = switches/usage, main, share, type="usage")],
  pwr_dts$digest[,list(switch_rate = switches/usage, main, share, type="power")],
  vMF_dts$digest[,list(switch_rate = switches/usage, main, share, type="peaktime")]
)
ggplot(digests) + facet_grid(type ~ .) + theme_bw() + aes(x=log10(switch_rate), fill=main, alpha=factor(floor(share*10))) + geom_bar(binwidth=0.1) + scale_alpha_manual(breaks=0:10, values=seq(0.5,1,length.out = 11))
ggplot(digests) + facet_grid(type ~ .) + theme_bw() + aes(x=log10(switch_rate), fill=main, alpha=factor(floor(share*10))) + geom_bar(binwidth=0.1, position="fill") + scale_alpha_manual(breaks=0:10, values=seq(0.5,1,length.out = 11))
```
\elandscape\newpage

\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="user categorization, by primary location type utilization"}
plot_ref <- ref[lifetime_dts$digest[, list(lifetime_main = main), keyby=user_id]][pwr_dts$digest[, list(pwr_main = main), keyby=user_id]][vMF_dts$digest[, list(peak_main = main), keyby=user_id]]
#thing[, list(user_id, mains=interaction(lifetime_main, pwr_main, peak_main))][,.N,by=mains]
ggplot(plot_ref[life > 365*24*60*60 & ul > 2 & ul < 20, list(y=log10(usage), x=log10(life), pwr_main, lifetime_main, peak_main)]) +facet_grid(pwr_main ~ lifetime_main) + aes(x=x,y=y, color=factor(peak_main)) + geom_point(alpha=0.3) + theme_bw()
```
\elandscape\newpage

>>>>

#lls <- lapply(8:11, function(k) movMF(pos, k, list(maxiter=10000, nruns=20)))
#bics <- lapply(lls, stats::BIC)
vmfmodel <- movMF(pos, 3, list(maxiter=10000, nruns=20))
temporal_cluster <- data.table(location_id = peaks[measure == "usage"]$location_id, vMFcluster = predict(vmfmodel, pos), key="location_id")
vmfrename <- peaks[measure == "usage"][temporal_cluster][,
  list(med=median(peak)),by=vMFcluster
][,vMFcluster, keyby=med][,list(vMFcluster, rename=factor(c("early","middle","late"), levels=c("early","middle","late"), ordered = T))]
temporal_cluster <- merge(temporal_cluster, vmfrename, by="vMFcluster")[,list(vMFcluster=rename), keyby=location_id]

#alt_temporal_cluster <- data.table(location_id = peaks[measure == "usage"]$location_id, vMFcluster = predict(lls[[1]], pos), key="location_id")

@@

```{r, echo=F, fig.cap="Login, Logout, Usage Peak Clusters"}
ggplot(peaks[temporal_cluster]) + theme_bw() + aes(x = peak, fill=vMFcluster) + facet_grid(measure ~ .) + geom_bar(binwidth=1) + labs(fill="von Mises-Fisher\n3-Sphere\nCluster\nPeak Time")
```

@@

Showing the power cluster breakdown
```{r}
pre_pwr_clust <- data.table(dcast(pwr_cluster, location_id ~ measure, value.var = "pwr_cat"))
pwr_multi_clusters <- setorder(pre_pwr_clust[,.N,by=list(login, logout, usage)], -N)
pwr_multi_clusters
reduced_pwr_clust <- pre_pwr_clust[, list(pwr_clust = factor(ifelse((login == logout) || (login == usage), login, logout), levels=c("lo","med","hi"), ordered=T)), keyby=location_id]
```
Probably worth having all-(low|med|high) clusters; the other cases all (with one exception) are two the same, the third off by one.

@@

Finally, we can look at the relative usage per user:
\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="Locations clustered excess usage per user."}
agg.dt <- initial_digest$raw[,list(log10_total_duration,log10_unique_users,log10_lifetime), keyby=location_id]
lg10lm <- lm(log10_total_duration-log10_unique_users ~ log10_lifetime, agg.dt)
agg.dt <- cbind(agg.dt, predict(lg10lm, agg.dt, interval="prediction", level=1/3))[,
  lifetime_cat := factor(ifelse(log10_total_duration-log10_unique_users < lwr, "low",
                  ifelse(log10_total_duration-log10_unique_users < upr, "mid", "high")), levels=c("low","mid","high"), ordered=T)
][reduced_pwr_clust][temporal_cluster]
p <- ggplot(agg.dt) + theme_bw() +
  aes(y = log10_total_duration-log10_unique_users, x = log10_lifetime, color = vMFcluster, size = pwr_clust, alpha = pwr_clust) +
  with(agg.dt, annotate("ribbon", x=log10_lifetime, ymax=upr, ymin=lwr, fill="lightgrey")) +
  geom_vline(xintercept=log10(365*24*60*60), color="grey") +
  annotate("text", x=log10(365*24*60*60)+c(-0.02,0.02), y=5.75, label=c("< 1 year","> 1 year"), angle=90, color="grey") +
  geom_point() +
#  geom_ribbon(aes(ymax=upr, ymin=lwr, y=NULL, color=NULL, size=NULL), alpha=0.2) +
  labs(
    x = expression(log[10]("lifetime")),
    y = expression(log[10]*bgroup("(",frac("total usage","# unique users"),")")),
    color = "von Mises-Fisher\nPeak Time\nCluster"
  ) + scale_size_manual("power\ncluster", limits=c("low","med","hi"), values=c(2,5,8)) +
  scale_alpha_manual("power\ncluster", limits=c("low","med","hi"), values=(2^0.5)/c(2,5,8)^0.5)
p
```
\elandscape\newpage

@@

categorizer <- function(src.dt, cat, ref) {
  rle_expr <- paste0("rle(",cat,")")
  key_expr <- c("user_id",cat)
  lvls <- levels(src.dt[[cat]])
  mx_expr <- parse(text=paste0("list(main=factor(",cat,"[which.max(prop)], ordered=T, levels=lvls),share=max(prop))"))
  rle_dt <- src.dt[, rle(as.numeric(eval(parse(text=cat)))), keyby=user_id]
  prop_ref <- src.dt[,list(prop=sum(logout-login)), keyby=key_expr][ref][,list(prop=prop/usage), keyby=key_expr]
  prop_lab <- prop_ref[, eval(mx_expr), by=user_id]
  chunks <- data.table(dcast(rle_dt, user_id ~ values, value.var="lengths", fun.aggregate = sum), key="user_id")
  ## cast data, maybe?
  digest <- rle_dt[, list(tot_visits = sum(lengths), mn_reps=mean(lengths), switches=.N), keyby=user_id][ref][prop_lab][chunks]
  list(rle_dt=rle_dt, prop_ref=prop_ref, prop_lab=prop_lab, digest=digest)
}

@@

\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="distribution of main location type usage"}
digests <- rbind(
  lifetime_dts$digest[,list(switch_rate = switches/usage, main, share, type="usage")],
  pwr_dts$digest[,list(switch_rate = switches/usage, main, share, type="power")],
  vMF_dts$digest[,list(switch_rate = switches/usage, main, share, type="peaktime")]
)
ggplot(digests) + facet_grid(type ~ .) + theme_bw() + aes(x=log10(switch_rate), fill=main, alpha=factor(floor(share*10))) + geom_bar(binwidth=0.1) + scale_alpha_manual(breaks=0:10, values=seq(0.5,1,length.out = 11))
```
\elandscape\newpage

\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="normalized distribution of main location type usage"}
ggplot(digests) + facet_grid(type ~ .) + theme_bw() + aes(x=log10(switch_rate), fill=main, alpha=factor(floor(share*10))) + geom_bar(binwidth=0.1, position="fill") + scale_alpha_manual(breaks=0:10, values=seq(0.5,1,length.out = 11))
```
\elandscape\newpage

\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="user categorization, by primary location type utilization"}
plot_ref <- ref[
  lifetime_dts$digest[, list(lifetime_main = main, lifetime_share = share), keyby=user_id]
][
  pwr_dts$digest[, list(pwr_main = main, pwr_share = share), keyby=user_id]
][
  vMF_dts$digest[, list(peak_main = main, peak_share = share), keyby=user_id]
]
user_src.dt <- plot_ref[between(life, 365*24*60*60, 3*365*24*60*60) & between(ul,3,20)]
#thing[, list(user_id, mains=interaction(lifetime_main, pwr_main, peak_main))][,.N,by=mains]
ggplot(user_src.dt[, list(y=log10(usage), x=log10(life), pwr_main, lifetime_main, peak_main, membership = lifetime_share * pwr_share * peak_share)]) +facet_grid(pwr_main ~ lifetime_main, labeller = function(var, val){
  if (var == "pwr_main") {
    ifelse(val == "lo", "smooth usage", ifelse(val=="med","some usage variability","highly variable usage"))
  } else {
    ifelse(val == "low", "low usage per time", ifelse(val=="mid","typical usage per time","high usage per time"))
  }
}) + aes(x=x,y=y, color=factor(peak_main), alpha = membership) + geom_point() + theme_bw() + labs(y="log(usage)", x="log(life)", color="usage\npeak")
```
\elandscape\newpage

```{r, echo=F}
# thing <- 
#   data.table(
#     x=lifetime_dts$digest[, list(pos = (0.5*`2`+1*`3`)/(`1`+`2`+`3`), main, share, user_id)]$pos,
#     y=pwr_dts$digest[, list(pos = (0.5*`2`+1*`3`)/(`1`+`2`+`3`), main, share, user_id)]$pos,
#     z=vMF_dts$digest[, list(pos = (0.5*`2`+1*`3`)/(`1`+`2`+`3`), main, share, user_id)]$pos,
#     key=c("x","y","z")
#   )
# uthing <- unique(thing)
# lenthing <- dim(uthing)[1]
# cloud(z ~ x*y, data=unique(thing)[sample.int(lenthing, 250)])
# censor.dt[user_id %in% user_src.dt$user_id, .N*60*60*24*30/(max(logout)-min(login)), by=user_id][V1>1]
```
